---
title: "Real Property Tax in Baltimore"
subtitle: "Flaws in the current SDAT assessments of land"
author: "Joshua Spokes"
---

```{r pull-city-data, cache = TRUE}
# Load necessary libraries
<<<<<<< HEAD
library(sf)
library(stringr)
library(dplyr)
library(tibble)
library(gt)
library(webshot2)
library(ggplot2)
library(tmap)
library(tmaptools)
library(OpenStreetMap)
library(units)
library(arcgislayers)
library(arcgis)
=======
library(httr)
library(jsonlite)
library(sf)
library(pbapply)
library(stringr)
library(dplyr)
library(NISTunits)
library(units)
>>>>>>> 0b4a764fe8345fd759afcbcbeb7aaa145b918ccc

# I am so angry that this works
baltimore <- st_read("https://geodata.baltimorecity.gov/egis/rest/services/CityView/Realproperty_OB/FeatureServer/0/query?returnGeometry=true&where=1=1&outFields=*&f=geojson")

baltimore_valid <- st_make_valid(baltimore)

mutate(baltimore_valid,
       total_assessed_value = CURRLAND + CURRIMPR + LANDEXMP + IMPREXMP,
       area_acres = set_units(st_area(geometry), acres),
       area_ft_sq = set_units(st_area(geometry), ft^2),
       valueperacre = total_assessed_value / area_acres,
<<<<<<< HEAD
       landvaluepsf = round((CURRLAND + LANDEXMP) / area_ft_sq, digits = 2)) -> baltimore_areas
```

```{r whole-neighborhoods}

filter(baltimore_areas,
       NEIGHBOR %in% c("BELAIR-EDISON                            ",
                       "ROLAND PARK                              ")
       ) |>
  select(FULLADDR,
         valueperacre,
         NEIGHBOR,
         total_assessed_value,
         area_acres) -> quicklook

st_drop_geometry(baltimore_areas) |>
group_by(NEIGHBOR) |>
  summarise(total_currvalue = sum(CURRLAND) + sum(CURRIMPR),
            ave_vpa_currland_only = (sum(CURRLAND) + sum(CURRIMPR)) / sum(area_acres),
            avg_vpa = sum(total_assessed_value) / sum(area_acres)) -> avg_neighborhood_vpa

write.csv(avg_neighborhood_vpa, "neighborhood_vpas.csv")
st_drop_geometry(baltimore_areas) -> baltimore_areas_no_geo
```

```{r testing}
baltimore_valid |>
  mutate(NEIGHBOR = trimws(NEIGHBOR)) |>
  filter(NEIGHBOR == "HAMPDEN") -> hampden

hampden |> 
  smallest_area(value = "CURRLAND", percent = 10) -> smalles_area_hampden

```

```{r initial analysis}
initial_analysis_sample = c("1783017A", "1783017", "1784040", "1784036",
                            "1784035", "1784034", "1784033", "1784032",
                            "1784031", "1784030", "1784029", "1784028",
                            "1784027", "1784026", "1784025", "1784024",
                            "1784023", "1784022", "1784021", "1784020",
                            "1784019", "1784018", "1784017", "1784016",
                            "1784061", "1784060", "1784059", "1784058",
                            "1784057", "1784056", "1784055", "1784054",
                            "1784053", "1784052")

filter(baltimore_areas_no_geo,
       PIN %in% initial_analysis_sample) |>
  drop_units() |>
  group_by(BLOCK) |>
  summarise("Value Per Acre ($/acre)" = sum(total_assessed_value) / sum(area_acres),
            "City Property Tax ($)" = sum(CITY_TAX),
            "Land Value PSF ($/sf)" = round(sum(CURRLAND) / sum(area_ft_sq), digits = 2)) |>
  t() |>
  as.data.frame() |>
  rownames_to_column() |>
  rename(" " = rowname,
         "Burger King/RoFo Block" = V1,
         "Patterson Bowling Block" = V2
         ) -> table_df

gt(table_df) |>
  fmt_number(decimals = 1,
             suffixing =  TRUE) |>
    cols_align(
    align = "center",  # Specify center alignment
    columns = everything()  # Apply to all columns
  ) |>
  gtsave("Images/initial_analysis_comparison.png")
=======
       landvaluepsf = (CURRLAND + LANDEXMP) / area_ft_sq) -> baltimore_areas
```

```{r initial analysis}
initial_analysis_sample = c("1783017A",
                            "1783017",
                            "1784040",
                            "1784036",
                            "1784035",
                            "1784034",
                            "1784033",
                            "1784032",
                            "1784031",
                            "1784030",
                            "1784029",
                            "1784028",
                            "1784027",
                            "1784026",
                            "1784025",
                            "1784024",
                            "1784023",
                            "1784022",
                            "1784021",
                            "1784020",
                            "1784019",
                            "1784018",
                            "1784017",
                            "1784016",
                            "1784061",
                            "1784060",
                            "1784059",
                            "1784058",
                            "1784057",
                            "1784056",
                            "1784055",
                            "1784054",
                            "1784053",
                            "1784052")

filter(baltimore_areas,
       PIN %in% initial_analysis_sample) -> initial_analysis_lots

filter(initial_analysis_lots,
       str_detect(PIN, "^1783")) -> bk_rofo_block

rofo_vpa <- sum(bk_rofo_block$total_assessed_value) / sum(bk_rofo_block$area_acres)
rofo_prop_tax <- sum(bk_rofo_block$CITY_TAX)
rofo_land_value_psf <- sum(bk_rofo_block$CURRLAND) / sum(bk_rofo_block$area_ft_sq)

filter(initial_analysis_lots,
       str_detect(PIN, "^1784")) |>
  mutate(land_psf = CURRLAND / area_ft_sq) -> patterson_block

patterson_vpa <- sum(patterson_block$total_assessed_value) / sum(patterson_block$area_acres)
patterson_prop_tax <- sum(patterson_block$CITY_TAX)
patterson_land_value_psf <- sum(patterson_block$CURRLAND) / sum(patterson_block$area_ft_sq)

patterson_vpa

>>>>>>> 0b4a764fe8345fd759afcbcbeb7aaa145b918ccc
```

```{r two-block-city}
filter(baltimore_areas,
<<<<<<< HEAD
       BLOCK %in% c("1783 ", "1784 ")) |>
  select(FULLADDR,
         total_assessed_value,
         area_ft_sq,
         area_acres,
         landvaluepsf,
         valueperacre
         ) ->  two_block_city

slice_max(two_block_city,
          landvaluepsf,
          n = 5
          ) |>
  st_buffer(dist = 2) -> highest_assessed_land

slice_min(two_block_city,
          landvaluepsf,
          n = 5
          ) |>
  arrange(desc(landvaluepsf)) -> lowest_assessed_land

tmap_mode("view")

two_block_city_map <- tm_basemap(leaflet::providers$Esri.WorldImagery) +
tm_shape(two_block_city) + 
   tm_fill(col = "gray", alpha = 0.2) +
tm_shape(lowest_assessed_land) +
  tm_borders(col = "red", lwd = 2) +
  tm_text("landvaluepsf", col = "white", bg.color = "black") +
tm_shape(highest_assessed_land) +
  tm_fill(col = "green", lwd = 2) +
  tm_text("landvaluepsf", col = "black", auto.placement = TRUE)

tmap_save(two_block_city_map, "Images/two_block_city.html")
```
=======
       BLOCK %in% c("1783 ", "1784 ")) -> two_block_city

slice_max(two_block_city, landvaluepsf, n = 5) |>
 arrange(desc(landvaluepsf)) |>
  select(FULLADDR, total_assessed_value, area_ft_sq, area_acres, landvaluepsf, valueperacre) -> highest_assessed_land

highest_assessed_land

slice_min(two_block_city, landvaluepsf, n = 5) |>
 arrange(desc(landvaluepsf)) |>
  select(FULLADDR, total_assessed_value, area_ft_sq, area_acres, landvaluepsf, valueperacre) -> lowest_assessed_land

lowest_assessed_land

library(tmap)

tmap_mode("view")

tm_basemap(leaflet::providers$OpenStreetMap) +
tm_shape(two_block_city) + 
  tm_polygons(border.col = "black") +
tm_shape(lowest_assessed_land) +
  tm_polygons(col = "red") +
tm_shape(highest_assessed_land) +
  tm_polygons(col = "green")

```



>>>>>>> 0b4a764fe8345fd759afcbcbeb7aaa145b918ccc
